### 2024-08-02 SFE_poster_DOE/conference_poster_6.tex
02.08

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% a0poster Portrait Poster
% LaTeX Template
% Version 1.0 (22/06/13)
%
% The a0poster class was created by:
% Gerlinde Kettl and Matthias Weiser (tex@kettl.de)
% 
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[a0,portrait]{a0poster}

\usepackage{multicol} % This is so we can have multiple columns of text side-by-side
\columnsep=100pt % This is the amount of white space between the columns in the poster
\columnseprule=3pt % This is the thickness of the black line between the columns in the poster

\usepackage[svgnames]{xcolor} % Specify colors by their 'svgnames', for a full list of all colors available see here: http://www.latextemplates.com/svgnames-colors

\usepackage{times} % Use the times font
%\usepackage{palatino} % Uncomment to use the Palatino font

\usepackage{graphicx} % Required for including images
\graphicspath{{figures/}} % Location of the graphics files
\usepackage{booktabs} % Top and bottom rules for table
\usepackage[font=small,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\usepackage{amsfonts, amsmath, amsthm, amssymb} % For math fonts, symbols and environments
\usepackage{wrapfig} % Allows wrapping text around tables and figures

\usepackage{tcolorbox}

\begin{document}
\pagecolor{white!90!black}
%----------------------------------------------------------------------------------------
%	POSTER HEADER 
%----------------------------------------------------------------------------------------

% The header is divided into two boxes:
% The first is 75% wide and houses the title, subtitle, names, university/organization and contact information
% The second is 25% wide and houses a logo for your university/organization or a photo of you
% The widths of these boxes can be easily edited to accommodate your content as you see fit

\begin{minipage}[b]{0.75\linewidth}
\veryHuge \color{NavyBlue} \textbf{Optimal design of experiment:} \color{Black}\\ % Title
\Huge\textit{supercritical fluid extraction case}\\[2cm] % Subtitle
\huge \textbf{Oliwer Sliczniuk \& Pekka Oinas}\\[0.5cm] % Author(s)
\huge School of Chemical Engineering, Aalto University, Finland\\[0.4cm] % University/organization
%\Large \texttt{john@LaTeXTemplates.com} --- 1 (000) 111 1111\\
\end{minipage}
%
\begin{minipage}[b]{0.25\linewidth}
\includegraphics[width=15cm]{Aalto_University_logo.png}\\ \\
\end{minipage}

%\vspace{0.5cm} % A bit of extra whitespace between the header and poster content
\noindent\rule{\textwidth}{3pt}
%\vspace{0.25cm} % A bit of extra whitespace between the header and poster content

%----------------------------------------------------------------------------------------

\begin{multicols}{2} % This is how many columns your poster will be broken into, a portrait poster is generally split into 2 columns

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

%\color{Navy} % Navy color for the abstract

%\begin{abstract}

%\end{abstract}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

%\color{Navy} % SaddleBrown color for the introduction

\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
	{\LARGE Introduction\\}

This study investigates the extraction of valuable components from biomass using supercritical carbon dioxide as a solvent. The interest is in essential oils extracted from chamomile flowers in cylindrical extractors operated in a semi-batch mode. The extraction process is described by a distributed-parameter model, and set of empirical correlations. This work aims \underline{to validate the empirical correlations by designing a new experiment} with dynamically changing operating conditions. %The process model includes empirical correlations, which describe the relation between the extraction kinetics parameters and the operating conditions. 
\end{tcolorbox}

%----------------------------------------------------------------------------------------
%	OBJECTIVES
%----------------------------------------------------------------------------------------

%\color{DarkSlateGray} % DarkSlateGray color for the rest of the content

%\section*{Main Objectives}

%\begin{enumerate}
%\item Lorem ipsum dolor sit amet, consectetur.
%\item Nullam at mi nisl. Vestibulum est purus, ultricies cursus volutpat sit amet, vestibulum eu.
%\item Praesent tortor libero, vulputate quis elementum a, iaculis.
%\item Phasellus a quam mauris, non varius mauris. Fusce tristique, enim tempor varius porta, elit purus commodo velit, pretium mattis ligula nisl nec ante.
%\item Ut adipiscing accumsan sapien, sit amet pretium.
%\item Estibulum est purus, ultricies cursus volutpat
%\item Nullam at mi nisl. Vestibulum est purus, ultricies cursus volutpat sit amet, vestibulum eu.
%\item Praesent tortor libero, vulputate quis elementum a, iaculis.
%\end{enumerate}

%----------------------------------------------------------------------------------------
%	MATERIALS AND METHODS
%----------------------------------------------------------------------------------------

%\section*{Materials and Methods}

%Fusce magna risus, molestie ut porttitor in, consectetur sed mi. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Pellentesque consectetur blandit pellentesque. Sed odio justo, viverra nec porttitor vel, lacinia a nunc. Suspendisse pulvinar euismod arcu, sit amet accumsan enim fermentum quis. In id mauris ut dui feugiat egestas. Vestibulum ac turpis lacinia nisl commodo sagittis eget sit amet sapien.

%------------------------------------------------

%\vspace{0.1cm}
%\section*{Process Model}
\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
	{\LARGE Process Model\\}

\begin{minipage}[b]{0.45\linewidth}
	\begin{itemize}
		\item One-dimensional
		\item Plug flow
		\item No pressure drop
		\item Uniform particle distribution
	\end{itemize}
\end{minipage}
\hfill
\begin{minipage}[b]{0.45\linewidth}
	\begin{itemize}
		\item Based on the two-film theory for a single component
		\item Peng-Robinson equation of state
		\item Decaying extraction kinetic
		\item Empirical correlations
	\end{itemize}
\end{minipage}

%\begin{minipage}[b]{0.5\linewidth}
%	\begin{itemize}
%		\item Based on the two-film theory for a single component
%		\item Peng-Robinson equation of state
%		\item Decaying extraction kinetic
%		\item Empirical correlations
%	\end{itemize}
%\end{minipage}

%The extraction process is represented by a \underline{distributed-parameter model} consisting of partial differential equations. This model has been validated within the operating conditions: temperatures between $30 - 40^\circ C$, pressures between $100 - 200$ bar, and mass flow rates between $3.33 - 6.67 \times 10^{-5}$ kg/s. The solute extraction from the solid to the fluid phase is described using the two-film theory for a single component. The physical properties of the fluid and pseudo-homogeneous phase are considered to be locally dependent on temperature and pressure and are calculated using the Peng-Robinson equation of state. By incorporating the enthalpy balance, the extraction kinetic becomes a function of operating conditions. Empirical correlations were developed to generalize the process model across the entire range of operating conditions.

\begin{align*}
	\dot{x} = \cfrac{d x}{d t} = 
	\begin{bmatrix}
		\frac{\partial c_f}{\partial t} \\
		\\
		\cfrac{\partial c_s}{\partial t} \\
		\\
		\cfrac{\partial \left(\rho_f h A_f\right)}{\partial t} \\
		\\
		\cfrac{d y}{d t}
	\end{bmatrix}
	=
\underbrace{\begin{bmatrix}
		-\frac{1}{\phi} \frac{\partial \left( c_f u\right)}{\partial z} + \frac{1-\phi}{\phi} \cfrac{\partial c_s}{\partial t} + \frac{1}{\phi} \frac{\partial}{\partial z} \left( D^M_e \frac{\partial c_f}{\partial z} \right) \\
		-\cfrac{ D_i^R \exp \left( \Upsilon \left( 1-\cfrac{ c_s }{c_{s0}} \right) \right) }{ \mu l^2 } \left( c_s  - \cfrac{\rho_s c_f }{ k_m \rho_f }  \right) \\
		- \cfrac{\partial \left( \rho_f h A_f {\color{red}v} \right)}{\partial z} + \cfrac{\partial \left(P A_f\right)}{\partial t} + \cfrac{\partial}{\partial z} \left( k \cfrac{\partial T}{\partial z} \right) \\
		\cfrac{F}{\rho_f} c_f \biggr\rvert_{z=L}
	\end{bmatrix}}_{G(x,t,\Theta;\Xi)}
\end{align*}

\begin{minipage}[h!]{0.5\linewidth}
	\begin{align*}
		c_f &- \text{Soluts concentration in the fluid phase}\\
		c_s &- \text{Soluts concentration in the solid phase}\\
		h~  &- \text{Enthalpy}\\
		y~ &- \text{Extraction yield}\\
		\rho_f &- \text{Density of fluid}\\
		A_f &- \text{Cross-section of the bed}\\
		u~~ &- \text{Darcy velocity}\\
		\phi~~ &- \text{Void fraction}
	\end{align*}
\end{minipage}
%
\begin{minipage}[h!]{0.5\linewidth}
	\begin{align*}
		D_e^M &- \text{Axial mass diffusivity}\\
		D_i^R &- \text{Internal mass diffusivity}\\
		\Upsilon~~ &- \text{Decaying factor}\\
		\rho_s~ &- \text{Bulky density of solid bed}\\
		~k_m &- \text{Partition factor}\\
		P~~ &- \text{Pressure}\\
		T~~ &- \text{Temperature}\\
		F~~ &- \text{Mass flow rate}
	\end{align*}
\end{minipage}

\begin{center}%\vspace{1cm}
	\includegraphics[width=0.49\linewidth]{Di_Re_F_1.png}
	\includegraphics[width=0.49\linewidth]{Gamma_Re_F_1.png}
	\captionof{figure}{Empirical correlations}
\end{center}%\vspace{1cm}

\end{tcolorbox}

\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
%------------------------------------------------
%\vspace{1cm}
%\section*{Model-based optimal design of experiment}
\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
	{\LARGE Model-based optimal design of experiment\\}
\end{tcolorbox}

\underline{Fisher information $\mathcal{F}$ measures the amount of information} that an observable random variable carries about an unknown parameter of a distribution that models the random variable. $\mathcal{F}$ is defined as Hessian of the log-likelihood function with respect to the parameters $\Theta$, given the covariance matrix $\Sigma$. %Based on the Cramer-Rao inequality, $\mathcal{F}$ can used to calculate the covariance matrices associated with maximum-likelihood estimates.

\begin{equation*}
	%\Sigma^{-1} \geq
	\mathcal{F}(t,\Theta; \Xi) = \frac{\partial y(t, \Theta; \Xi)}{\partial \Theta} \Sigma \frac{\partial y(t, \Theta; \Xi)}{\partial \Theta^\top}
\end{equation*}

The optimal design of experiments is a concept that refers to planning an experiment, which allow parameters to be estimated without bias and with minimum variance. The D-optimality criterion is selected as the objective function, which leads to \underline{the minimisation of the volume of the ellipsoidal confidence region of parameter estimates} given the experimental condition $\Xi$.

\begin{equation*}
	\begin{aligned} 
		&\Xi^* &= \arg &\min_{ T^{in},~F\in~\Xi}\quad\int_{t_0}^{t_f} - \ln \det \mathcal{F}(t,\Theta; \Xi)~dt  \\
		&\text{subject to}
		& \dot{x} = ~&G~~(x,t,\Theta;\Xi) \\
		&& T^{0} = ~&T^{in}(t=0) \\
		&& 30^\circ C \leq ~&T^{in}(t) ~ \leq 40^\circ C \\
		&& 3.33 \cdot 10^{-5}~\text{kg/s} \leq ~&F~~(t)~ \leq 6.67 \cdot 10^{-5}~\text{kg/s}\\
		&& 100~\text{bar} \leq ~&P~~(t)~ \leq 200~\text{bar} \\
	\end{aligned}
\end{equation*}

\underline{This work aims to design a dynamic experiment to improve precision of correlation for $D_i^R$}. The system is assumed to run for 300 min, with sampling time of 10 min. The set of decision variables consists of $T^{in}$ and $F$, and each of them is manipulated every 15 min. Each of five analysed cases assumes that P is kept constant at 100, 125, 150, 175 and 200 bar, respectively.
\end{tcolorbox}
%----------------------------------------------------------------------------------------
%	RESULTS 
%----------------------------------------------------------------------------------------
\begin{tcolorbox}[width=\linewidth, boxrule=0mm, sharp corners=all, colback=white]
	{\LARGE Results\\}

	To identify the global solution, the optimization problem is solved multiple times, each starting from a random initial solution. Figure 2 compares the initial and final values of the cost function across multiple optimization runs for different cases of pressure value.

	\begin{center}%\vspace{1cm}
		\includegraphics[width=\linewidth]{scatter.png}
		\captionof{figure}{Values of the objective function}
	\end{center}%\vspace{1cm}

	The optimal profiles of the inlet temperature and flow rates for each case are showed in Figure 3

	\begin{center}%\vspace{1cm}
		\includegraphics[width=0.49\linewidth]{Profile_T.png}
		\includegraphics[width=0.49\linewidth]{Profile_F.png}
		\captionof{figure}{Optimal profiles}
	\end{center}%\vspace{1cm}

\end{tcolorbox}

%----------------------------------------------------------------------------------------
%	CONCLUSIONS
%----------------------------------------------------------------------------------------

\color{SaddleBrown} % SaddleBrown color for the conclusions to make them stand out

\section*{Conclusions}

\begin{itemize}
\item Pellentesque eget orci eros. Fusce ultricies, tellus et pellentesque fringilla, ante massa luctus libero, quis tristique purus urna nec nibh. Phasellus fermentum rutrum elementum. Nam quis justo lectus.
\item Vestibulum sem ante, hendrerit a gravida ac, blandit quis magna.
\item Donec sem metus, facilisis at condimentum eget, vehicula ut massa. Morbi consequat, diam sed convallis tincidunt, arcu nunc.
\item Nunc at convallis urna. isus ante. Pellentesque condimentum dui. Etiam sagittis purus non tellus tempor volutpat. Donec et dui non massa tristique adipiscing.
\end{itemize}

\color{DarkSlateGray} % Set the color back to DarkSlateGray for the rest of the content

%----------------------------------------------------------------------------------------
%	FORTHCOMING RESEARCH
%----------------------------------------------------------------------------------------

\section*{Forthcoming Research}

Vivamus molestie, risus tempor vehicula mattis, libero arcu volutpat purus, sed blandit sem nibh eget turpis. Maecenas rutrum dui blandit lorem vulputate gravida. Praesent venenatis mi vel lorem tempor at varius diam sagittis. Nam eu leo id turpis interdum luctus a sed augue. Nam tellus.

 %----------------------------------------------------------------------------------------
%	REFERENCES
%----------------------------------------------------------------------------------------

\nocite{*} % Print all references regardless of whether they were cited in the poster or not
\bibliographystyle{plain} % Plain referencing style
\bibliography{sample} % Use the example bibliography file sample.bib

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\section*{Acknowledgements}

Etiam fermentum, arcu ut gravida fringilla, dolor arcu laoreet justo, ut imperdiet urna arcu a arcu. Donec nec ante a dui tempus consectetur. Cras nisi turpis, dapibus sit amet mattis sed, laoreet.

%----------------------------------------------------------------------------------------

\end{multicols}
\end{document}